---
title: "HW 12"
author: "Erika Sklaver"
date: "5/7/2017"
output: html_document
---

## Question 1 
\
\
\
\
\
\
\
\
\
\
\
\

The probability that a draw from the true sampling distribution will fall into the critical region is: 

```{r}
1-pnorm(qnorm(.99), 1,1)
```


This means that we will reject the null hypothesis $9.236%$ of the time.

```{r}
prob = rep(0,10)

for (i in 0:10){
  prob[i] = 1-pnorm(qnorm(.99),i,1)
}

plot(prob, xlab = "Mean Values", ylab = "Probability of Rejecting the Null")
```

## Question 2 

```{r}
sample_mean = .1
null_mean = 0 
var = 1
z_alpha = qnorm(.99)

probas = rep(NULL, 20)

for(N in 2:20){
 
accepted = rep(NULL, 1000)  
  for(i in 1:1000){
    data = rnorm(N, sample_mean, sqrt(var))
    u_hat = mean(data)
    z_score = (u_hat - null_mean) * sqrt(N) * (sqrt(var))
    
    if(z_score>=z_alpha){
      ##reject null
      accepted[i] = 1
    }else{
      ##don't reject null 
      accepted[i] = 0
    }
  }
  
  probas[N] = mean(accepted, na.rm = TRUE) 
}

plot(probas)


probs = rep(NULL, 20)
 

for(N in 2:20){
  
  accepted = rep(NULL, 1000)  
  for(i in 1:1000){
    data = rnorm(N, sample_mean, sqrt(var))
    u_hat = mean(data)
    var_hat = var(data)
    z_score = (u_hat - null_mean) * sqrt(N) * (sqrt(var_hat))
    
    if(z_score>=z_alpha){
      ##reject null
      accepted[i] = 1
    }else{
      ##don't reject null 
      accepted[i] = 0
    }
  }
  
  probs[N] = mean(accepted, na.rm = TRUE) 
 
}

plot(probas, col = "blue", ylim = c(0,.2), xlab = "N", ylab = "Probability")
points(probs, col = "red")

legend(x = "topright", legend = c("Standard Deviation = 1", "Sample Standard Deviation"),
       col = c("blue","red"), pch = 1)
```

When the standard deviation is 1 each time (blue), there is less variability in the distribution. This is why the probability of rejecting the null hypothesis is a bit lower than when calculating the standard deviation each time. 