---
title: "HW 13"
author: "Erika Sklaver"
date: "5/11/2017"
output: html_document
---

##Question 1 

#(a)
When N=20 

```{r}

N = 20 
theta = 0.2 
probs = rep(0,20)

for(k in 0:N){
  
  x1_vec = rep(0,k)
  for(x1 in 0:k){
    
    sum_up = 0;
    for(x2 in 0:k){
      sum_up = sum_up + choose(N, x1)*choose(N,x2)*choose(N-x1,k-x1)*choose(N-x2, N-k)*(1/choose(2*N-x1-x2,N-x1))*theta^(x1+x2)*(1-theta)^(2*N-x1-x2)
    }
    x1_vec[x1]=sum_up 
  }
  
  probs[k]= sum(x1_vec)
}

plot(probs, col = "red", ylim = c(0,.4), xlab = "K", ylab = "Probability")
lines(probs, col = "red")

##theta = .4 
theta = 0.4 
probs = rep(0,20)

for(k in 0:N){
  
  x1_vec = rep(0,k)
  for(x1 in 0:k){
    
    sum_up = 0;
    for(x2 in 0:k){
      sum_up = sum_up + choose(N, x1)*choose(N,x2)*choose(N-x1,k-x1)*choose(N-x2, N-k)*(1/choose(2*N-x1-x2,N-x1))*theta^(x1+x2)*(1-theta)^(2*N-x1-x2)
    }
    x1_vec[x1]=sum_up 
  }
  
  probs[k]= sum(x1_vec)
}

points(probs, col = "orange")
lines(probs, col = "orange")

##theta = .6 
theta = 0.6 
probs = rep(0,20)

for(k in 0:N){
  
  x1_vec = rep(0,k)
  for(x1 in 0:k){
    
    sum_up = 0;
    for(x2 in 0:k){
      sum_up = sum_up + choose(N, x1)*choose(N,x2)*choose(N-x1,k-x1)*choose(N-x2, N-k)*(1/choose(2*N-x1-x2,N-x1))*theta^(x1+x2)*(1-theta)^(2*N-x1-x2)
    }
    x1_vec[x1]=sum_up 
  }
  
  probs[k]= sum(x1_vec)
}

points(probs, col = "green")
lines(probs, col = "green")

##theta = .8 
theta = 0.8 
probs = rep(0,20)

for(k in 0:N){
  
  x1_vec = rep(0,k)
  for(x1 in 0:k){
    
    sum_up = 0;
    for(x2 in 0:k){
      sum_up = sum_up + choose(N, x1)*choose(N,x2)*choose(N-x1,k-x1)*choose(N-x2, N-k)*(1/choose(2*N-x1-x2,N-x1))*theta^(x1+x2)*(1-theta)^(2*N-x1-x2)
    }
    x1_vec[x1]=sum_up 
  }
  
  probs[k]= sum(x1_vec)
}

points(probs, col = "blue")
lines(probs, col = "blue")

legend(x = "topleft", legend = c("theta = 0.2", "theta = 0.4", "theta = 0.6", "theta = 0.8"),
       col = c("red", "orange", "green", "blue"), pch = 1)

```

When N = 50 

```{r}

N = 50 
theta = 0.2 
probs = rep(0,50)

for(k in 0:N){
  
  x1_vec = rep(0,k)
  for(x1 in 0:k){
    
    sum_up = 0;
    for(x2 in 0:k){
      sum_up = sum_up + choose(N, x1)*choose(N,x2)*choose(N-x1,k-x1)*choose(N-x2, N-k)*(1/choose(2*N-x1-x2,N-x1))*theta^(x1+x2)*(1-theta)^(2*N-x1-x2)
    }
    x1_vec[x1]=sum_up 
  }
  
  probs[k]= sum(x1_vec)
}

plot(probs, col = "red", ylim = c(0,.4), xlab = "K", ylab = "Probability")
lines(probs, col = "red")

##theta = .4 
theta = 0.4 
probs = rep(0,50)

for(k in 0:N){
  
  x1_vec = rep(0,k)
  for(x1 in 0:k){
    
    sum_up = 0;
    for(x2 in 0:k){
      sum_up = sum_up + choose(N, x1)*choose(N,x2)*choose(N-x1,k-x1)*choose(N-x2, N-k)*(1/choose(2*N-x1-x2,N-x1))*theta^(x1+x2)*(1-theta)^(2*N-x1-x2)
    }
    x1_vec[x1]=sum_up 
  }
  
  probs[k]= sum(x1_vec)
}

points(probs, col = "orange")
lines(probs, col = "orange")

##theta = .6 
theta = 0.6 
probs = rep(0,50)

for(k in 0:N){
  
  x1_vec = rep(0,k)
  for(x1 in 0:k){
    
    sum_up = 0;
    for(x2 in 0:k){
      sum_up = sum_up + choose(N, x1)*choose(N,x2)*choose(N-x1,k-x1)*choose(N-x2, N-k)*(1/choose(2*N-x1-x2,N-x1))*theta^(x1+x2)*(1-theta)^(2*N-x1-x2)
    }
    x1_vec[x1]=sum_up 
  }
  
  probs[k]= sum(x1_vec)
}

points(probs, col = "green")
lines(probs, col = "green")

##theta = .8 
theta = 0.8 
probs = rep(0,50)

for(k in 0:N){
  
  x1_vec = rep(0,k)
  for(x1 in 0:k){
    
    sum_up = 0;
    for(x2 in 0:k){
      sum_up = sum_up + choose(N, x1)*choose(N,x2)*choose(N-x1,k-x1)*choose(N-x2, N-k)*(1/choose(2*N-x1-x2,N-x1))*theta^(x1+x2)*(1-theta)^(2*N-x1-x2)
    }
    x1_vec[x1]=sum_up 
  }
  
  probs[k]= sum(x1_vec)
}

points(probs, col = "blue")
lines(probs, col = "blue")

legend(x = "topleft", legend = c("theta = 0.2", "theta = 0.4", "theta = 0.6", "theta = 0.8"),
       col = c("red", "orange", "green", "blue"), pch = 1)

```

# (b)

When N=4 

```{r}
# 4 cups 
N=4
KcN4 = qhyper(.95, N, N, N)

theta = 0.5

probs = rep(0, KcN4+1)

for (k in 0:KcN4){
  
  x1_vec = rep(0,k+1)
  for (x1 in 0:k){
    
    sum_up = 0
    for(x2 in 0:k){
      sum_up = sum_up + choose(N, x1)*choose(N,x2)*choose(N-x1,k-x1)*choose(N-x2, N-k)*(1/choose(2*N-x1-x2,N-x1))*theta^(x1+x2)*(1-theta)^(2*N-x1-x2)
    }
    x1_vec[x1+1]= sum_up
  }
  probs[k+1]=sum(x1_vec)
}

beta = sum(probs)
power4 = 1-beta
```
$k_c(N) = 4$ and the probability that she finds herself in the critical region is $0.3017299$


When N=10

```{r}
#10 cups 
N=10
KcN10 = qhyper(.95, N, N, N)

theta = 0.5

probs = rep(0, KcN10+1)

for (k in 0:KcN10){
  
  x1_vec = rep(0,k+1)
  for (x1 in 0:k){
    
    sum_up = 0
    for(x2 in 0:k){
      sum_up = sum_up + choose(N, x1)*choose(N,x2)*choose(N-x1,k-x1)*choose(N-x2, N-k)*(1/choose(2*N-x1-x2,N-x1))*theta^(x1+x2)*(1-theta)^(2*N-x1-x2)
    }
    x1_vec[x1+1]= sum_up
  }
  probs[k+1]=sum(x1_vec)
}

beta = sum(probs)
power10 = 1-beta


```
$k_c(N) = 8$ and the probability that she finds herself in the critical region is $0.5708551$


When N=20 

```{r}
#20 cups 
N=20
KcN20 = qhyper(.95, N, N, N)

theta = 0.5

probs = rep(0, KcN20+1)

for (k in 0:KcN20){
  
  x1_vec = rep(0,k+1)
  for (x1 in 0:k){
    
    sum_up = 0
    for(x2 in 0:k){
      sum_up = sum_up + choose(N, x1)*choose(N,x2)*choose(N-x1,k-x1)*choose(N-x2, N-k)*(1/choose(2*N-x1-x2,N-x1))*theta^(x1+x2)*(1-theta)^(2*N-x1-x2)
    }
    x1_vec[x1+1]= sum_up
  }
  probs[k+1]=sum(x1_vec)
}

beta = sum(probs)
power20 = 1-beta

```
$k_c(N) = 14$ and the probability that she finds herself in the critical region is $0.8806994$


## Question 2 

```{r}
##power of test when N=5
N = 5
null_p = 0.5
real_p = 0.6
alpha = .01

z_alpha_left = qbinom(p = alpha/2, size = N, prob = null_p)
z_alpha_right = qbinom(p = 1-alpha/2, size = N, prob = null_p)
beta = pbinom(z_alpha_right-1,size = N, prob = real_p) - pbinom(z_alpha_left+1,size = N, prob = real_p)

power5 = 1-beta
```
Power of the test when N=5 is: $0.1648$

```{r}
##power of test when N=10
N = 10
null_p = 0.5
real_p = 0.6
alpha = .01

z_alpha_left = qbinom(p = alpha/2, size = N, prob = null_p)
z_alpha_right = qbinom(p = 1-alpha/2, size = N, prob = null_p)
beta = pbinom(z_alpha_right-1,size = N, prob = real_p) - pbinom(z_alpha_left+1,size = N, prob = real_p)

power10 = 1-beta
```
Power of the test when N=10 is: $0.05865196$

```{r}
##power of test when N=20
N = 20
null_p = 0.5
real_p = 0.6
alpha = .01

z_alpha_left = qbinom(p = alpha/2, size = N, prob = null_p)
z_alpha_right = qbinom(p = 1-alpha/2, size = N, prob = null_p)
beta = pbinom(z_alpha_right-1,size = N, prob = real_p) - pbinom(z_alpha_left+1,size = N, prob = real_p)

power20 = 1-beta
```
Power of the test when N=20 is: $0.05256348$

```{r}
##power of test when N=100
N = 100
null_p = 0.5
real_p = 0.6
alpha = .01

z_alpha_left = qbinom(p = alpha/2, size = N, prob = null_p)
z_alpha_right = qbinom(p = 1-alpha/2, size = N, prob = null_p)
beta = pbinom(z_alpha_right-1,size = N, prob = real_p) - pbinom(z_alpha_left+1,size = N, prob = real_p)

power100 = 1-beta 
```
Power of the test when N=100 is: $0.3068171$

## Question 3 

```{r}
N = 25
stdv = 1
a = 0.05

x <- seq(-4, 4, length=100)

plot(x,dt(x, df = N-1), type='l', lty=1, xlab='', ylab='', ylim = c(0, .4))
lines(x,dnorm(x, mean = 1, sd= stdv), type='l', lty=1, xlab='', ylab='', col = "red")

legend(x = "topleft", legend = c("Null Distribution", "Sampling Distribution"),
       col = c("black","red"), lty=c(1,1), cex=.65)

crit_val = qt(1-a, df = N-1)
beta = pnorm(crit_val, mean = 1, sd = stdv)
power = 1-beta 
```

The power of the test is $0.2385787$

```{r}
alpha = seq(from = 0.01, to = .99, by = .01)

#N=2
N = 2 
power = rep(0, length(alpha))
for (i in 1:length(alpha)){
  crit_val = qt(1-alpha[i], df = N-1)
  beta = pnorm(crit_val, mean = 1, sd = stdv)
  power[i] = 1-beta
}
plot(alpha, power, col = "red", type = 'l')

#N=5
N = 5 
power = rep(0, length(alpha))
for (i in 1:length(alpha)){
  crit_val = qt(1-alpha[i], df = N-1)
  beta = pnorm(crit_val, mean = 1, sd = stdv)
  power[i] = 1-beta
}
points(alpha, power, col = "orange", type = 'l')

#N=10
N = 10 
power = rep(0, length(alpha))
for (i in 1:length(alpha)){
  crit_val = qt(1-alpha[i], df = N-1)
  beta = pnorm(crit_val, mean = 1, sd = stdv)
  power[i] = 1-beta
}
points(alpha, power, col = "green", type = 'l')

#N=30

N = 30 
power = rep(0, length(alpha))
for (i in 1:length(alpha)){
  crit_val = qt(1-alpha[i], df = N-1)
  beta = pnorm(crit_val, mean = 1, sd = stdv)
  power[i] = 1-beta
}
points(alpha, power, col = "blue", type = 'l')

#N=100

N = 100 
power = rep(0, length(alpha))
for (i in 1:length(alpha)){
  crit_val = qt(1-alpha[i], df = N-1)
  beta = pnorm(crit_val, mean = 1, sd = stdv)
  power[i] = 1-beta
}
points(alpha, power, col = "purple", type = 'l')

legend(x = "topleft", legend = c("N=2", "N=5", "N=10", "N=30", "N=100"),
       col = c("red", "orange", "green", "blue", "purple"), lty=c(1,1))

```


```{r}
mu = seq(from = 0, to = 5, by = .05)
N=100

#alpha = .01 
a = .01 
power = rep(0, length(mu))
for (i in 1:length(mu)){
  crit_val = qt(1-a, df = N-1)
  beta = pnorm(crit_val, mean = mu[i], sd = stdv)
  power[i] = 1-beta
}
plot(mu, power, col = "red", type = 'l', xlab = "means")

#alpha = .03
a = .03 
power = rep(0, length(mu))
for (i in 1:length(mu)){
  crit_val = qt(1-a, df = N-1)
  beta = pnorm(crit_val, mean = mu[i], sd = stdv)
  power[i] = 1-beta
}
points(mu, power, col = "orange", type = 'l')

#alpha = .05
a = .05 
power = rep(0, length(mu))
for (i in 1:length(mu)){
  crit_val = qt(1-a, df = N-1)
  beta = pnorm(crit_val, mean = mu[i], sd = stdv)
  power[i] = 1-beta
}
points(mu, power, col = "green", type = 'l')

#alpha = .08
a = .08 
power = rep(0, length(mu))
for (i in 1:length(mu)){
  crit_val = qt(1-a, df = N-1)
  beta = pnorm(crit_val, mean = mu[i], sd = stdv)
  power[i] = 1-beta
}
points(mu, power, col = "blue", type = 'l')

#alpha = .1
a = .1 
power = rep(0, length(mu))
for (i in 1:length(mu)){
  crit_val = qt(1-a, df = N-1)
  beta = pnorm(crit_val, mean = mu[i], sd = stdv)
  power[i] = 1-beta
}
points(mu, power, col = "purple", type = 'l')

legend(x = "topleft", legend = c("alpha = .01", "alpha = .03", "alpha = .05", "alpha = .07", "alpha = .1"),
       col = c("red", "orange", "green", "blue", "purple"), lty=c(1,1), cex = .9)

```